---
title: "SALT Survey Summary Report"
author: "SALT Management System"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc-depth: 2
    theme: cosmo
  pdf:
    toc: true
    documentclass: article
  docx:
    toc: true
execute:
  echo: false
  warning: false
  message: false
---

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
library(DT)

# Load data
data_long <- read.csv("data_long.csv")
data_wide <- read.csv("data_wide.csv")

# Set up knitr options
knitr::opts_chunk$set(fig.width = 10, fig.height = 6)
```

# Executive Summary

This report provides a comprehensive summary of survey responses collected through the SALT system as of `r format(Sys.Date(), "%B %d, %Y")`.

## Key Metrics

```{r}
total_surveys <- length(unique(data_long$survey_id))
total_participants <- length(unique(data_long$participant_id))
total_facilities <- length(unique(data_wide$meta_facility_name))
date_range <- range(as.Date(data_wide$meta_completed_at))

summary_table <- data.frame(
  Metric = c("Total Surveys", "Unique Participants", "Total Facilities", "Date Range"),
  Value = c(
    total_surveys,
    total_participants,
    total_facilities,
    paste(date_range[1], "to", date_range[2])
  )
)

kable(summary_table, caption = "Survey Statistics Overview")
```

# Response Analysis

## Variable Summary

```{r}
variable_summary <- data_long %>%
  group_by(variable) %>%
  summarise(
    n_responses = n(),
    n_unique = n_distinct(participant_id),
    missing_pct = round(mean(is.na(numeric_value) & is.na(text_value)) * 100, 1),
    mean_numeric = round(mean(numeric_value, na.rm = TRUE), 2)
  ) %>%
  arrange(desc(n_responses))

kable(head(variable_summary, 20),
      caption = "Top 20 Variables by Response Count",
      col.names = c("Variable", "Total Responses", "Unique Participants", "Missing %", "Mean (if numeric)"))
```

## Response Distribution

```{r, fig.cap="Response Counts for Top 10 Variables"}
top_variables <- variable_summary$variable[1:10]

data_long %>%
  filter(variable %in% top_variables) %>%
  group_by(variable) %>%
  summarise(count = n()) %>%
  ggplot(aes(x = reorder(variable, count), y = count, fill = variable)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  theme_minimal() +
  labs(title = "Response Counts for Top 10 Variables",
       x = "Variable",
       y = "Number of Responses") +
  guides(fill = "none") +
  theme(text = element_text(size = 12))
```

# Facility Analysis

## Surveys by Facility

```{r}
facility_summary <- data_wide %>%
  group_by(meta_facility_name) %>%
  summarise(
    n_surveys = n(),
    n_participants = n_distinct(participant_id),
    first_survey = min(as.Date(meta_completed_at)),
    last_survey = max(as.Date(meta_completed_at))
  ) %>%
  arrange(desc(n_surveys))

kable(facility_summary,
      caption = "Survey Activity by Facility",
      col.names = c("Facility", "Surveys", "Participants", "First Survey", "Last Survey"))
```

## Facility Performance

```{r, fig.cap="Survey Distribution Across Facilities"}
data_wide %>%
  group_by(meta_facility_name) %>%
  summarise(count = n()) %>%
  ggplot(aes(x = reorder(meta_facility_name, count), y = count, fill = meta_facility_name)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  theme_minimal() +
  labs(title = "Survey Distribution Across Facilities",
       x = "Facility",
       y = "Number of Surveys") +
  guides(fill = "none") +
  theme(text = element_text(size = 12))
```

# Temporal Trends

## Daily Survey Completions

```{r, fig.cap="Daily Survey Completion Trend"}
daily_completions <- data_wide %>%
  mutate(date = as.Date(meta_completed_at)) %>%
  group_by(date) %>%
  summarise(count = n()) %>%
  complete(date = seq.Date(min(date), max(date), by = "day"), fill = list(count = 0))

ggplot(daily_completions, aes(x = date, y = count)) +
  geom_line(color = "#3498db", size = 1) +
  geom_point(color = "#2c3e50", size = 2) +
  geom_smooth(method = "loess", se = TRUE, alpha = 0.2) +
  theme_minimal() +
  labs(title = "Daily Survey Completions Over Time",
       x = "Date",
       y = "Number of Surveys") +
  theme(text = element_text(size = 12))
```

## Weekly Patterns

```{r, fig.cap="Survey Completions by Day of Week"}
data_wide %>%
  mutate(
    date = as.Date(meta_completed_at),
    weekday = weekdays(date, abbreviate = TRUE),
    weekday = factor(weekday, levels = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"))
  ) %>%
  group_by(weekday) %>%
  summarise(count = n()) %>%
  ggplot(aes(x = weekday, y = count, fill = weekday)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Survey Completions by Day of Week",
       x = "Day of Week",
       y = "Number of Surveys") +
  guides(fill = "none") +
  theme(text = element_text(size = 12))
```

# Data Quality

## Completeness Analysis

```{r}
completeness_by_section <- data_long %>%
  mutate(section = sub("_.*", "", variable)) %>%
  group_by(section) %>%
  summarise(
    n_variables = n_distinct(variable),
    total_responses = n(),
    complete_responses = sum(!is.na(numeric_value) | !is.na(text_value)),
    completeness_pct = round((complete_responses / total_responses) * 100, 1)
  ) %>%
  arrange(desc(completeness_pct))

kable(head(completeness_by_section, 15),
      caption = "Data Completeness by Variable Section",
      col.names = c("Section", "Variables", "Total Responses", "Complete", "Completeness %"))
```

## Missing Data Patterns

```{r, fig.cap="Completeness Rate by Variable Section"}
completeness_by_section %>%
  head(15) %>%
  ggplot(aes(x = reorder(section, completeness_pct), y = completeness_pct, fill = completeness_pct)) +
  geom_bar(stat = "identity") +
  scale_fill_gradient(low = "#e74c3c", high = "#27ae60") +
  coord_flip() +
  theme_minimal() +
  labs(title = "Data Completeness by Variable Section",
       x = "Section",
       y = "Completeness (%)") +
  guides(fill = "none") +
  theme(text = element_text(size = 12))
```

---

*Report generated on `r format(Sys.Date(), "%B %d, %Y")` at `r format(Sys.time(), "%I:%M %p")`*