---
title: "Data Quality Report"
author: "SALT report builder"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc-depth: 3
    theme: journal
    embed-resources: true
    format-links: false
  pdf:
    toc: true
    documentclass: article
  docx:
    toc: true
execute:
  echo: false
  warning: false
  message: false
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(scales)

knitr::opts_chunk$set(fig.width = 9, fig.height = 5)

# ---- Load per the data contract -------------------------------------------
dl <- read.csv("data_long.csv", stringsAsFactors = FALSE)
dl$numeric_value <- suppressWarnings(as.numeric(dl$numeric_value))

# Build atomic wide frames from the documented long format. survey_id is the
# per-row key; first() guards against any accidental duplicate survey x var rows.
wn <- dl %>%
  select(survey_id, variable, numeric_value) %>%
  pivot_wider(names_from = variable, values_from = numeric_value,
              values_fn = dplyr::first)
wt <- dl %>%
  select(survey_id, variable, text_value) %>%
  pivot_wider(names_from = variable, values_from = text_value,
              values_fn = dplyr::first)

N <- nrow(wn)

# Safe accessor: returns an all-NA vector if a column is absent.
gv <- function(df, name) if (name %in% names(df)) df[[name]] else rep(NA, nrow(df))

# Logical helper that treats NA as FALSE (so counts are exact, not NA).
is_true <- function(x) !is.na(x) & x

# Collector for the findings summary.
issues <- tibble(Check = character(), Severity = character(), Finding = character())
add_issue <- function(check, severity, finding) {
  issues <<- add_row(issues, Check = check, Severity = severity, Finding = finding)
}

kable_if <- function(df, ...) {
  if (nrow(df) > 0) knitr::kable(df, ...) else cat("None detected.\n")
}
```

# Scope and approach

This report screens the **Short MSM Survey** (survey id 2, `r N` completed records) for
data errors: impossible values, range violations, implausible category frequencies, and
internal contradictions between related variables. It is a data-integrity check only; it does
not assess recruitment or RDS performance.

The screen is deliberately **conservative**. Skewed-but-real distributions are common in
key-population surveys, so a value is flagged only when it is outside a near-impossible
bound, sits on a known sentinel/ceiling, or contradicts another field. Naive interquartile
flagging is avoided.

```{r provenance-note, results='asis'}
dm <- unique(na.omit(gv(wt, "device_model")))
av <- unique(na.omit(gv(wt, "device_app_version")))
lg <- unique(na.omit(gv(wt, "meta_language")))
flag_synth <- any(grepl("fake", dm, ignore.case = TRUE)) ||
              any(grepl("seed", av, ignore.case = TRUE))
if (flag_synth) {
  add_issue("Provenance", "Critical",
            "Records carry synthetic markers (FakeData/seed device + single language); treat findings as demonstration, not field data.")
  cat(sprintf(
    "::: callout-important\nThis export appears to be **synthetic / seed data**: device_model = %s, app_version = %s, language = %s for all records. The checks below run normally, but the findings describe generated data, not real fieldwork.\n:::\n",
    paste(dm, collapse = ", "), paste(av, collapse = ", "), paste(lg, collapse = ", ")))
}
```

# Methods

Data are read from the contract files `data_long.csv` (one row per survey × variable) and
pivoted to one row per survey for numeric values (`wn`, option indices / numbers) and text
values (`wt`, labels). Cleaning and coding decisions:

- **Numeric coercion.** `numeric_value` is parsed to numeric; non-numeric becomes `NA`.
- **Skip logic vs missingness.** Questions hidden by skip rules (e.g. lab tests asked only when
  the HIV rapid test is positive; PrEP items asked only when PrEP is known) are *structurally*
  missing and are not counted as data-quality gaps. Skip integrity is itself checked below.
- **Sentinels / ceilings.** Counts that pile up exactly on a validation ceiling (e.g. 95 partners,
  99 violence events) are treated as suspected caps/sentinels, not as true magnitudes.
- **Labels.** Categorical logic uses the English text labels from `wt`; for HIV status the rapid
  result text ("positive"/"negative") and the confirmatory label are used directly.
- **Viral-load suppression** is defined as < 1000 copies/mL for the self-report-vs-lab check.
- **Aggregates only.** No per-respondent rows are printed; only counts, summaries, and the
  distinct offending values are shown.

A data-contract note: the export contains `coupon_issued_1/2/3` columns that are **not present in
the data dictionary** (which documents only `coupon_issued_count`). This is a documentation gap,
not a value error, but it means downstream code cannot rely on the dictionary alone.

# Numeric range and outlier checks

Each numeric variable is checked against a near-impossible bound. Histograms and the distinct
offending values are shown only where something is flagged.

```{r numeric-bounds}
bounds <- tribble(
  ~var,                 ~lower, ~upper, ~note,
  "q_tutage",               15,     89, "Eligibility range 15-89",
  "q_dgbehav",               1,    500, "Network: men known",
  "q_dgage",                 0,    500, "Network: aged 18+",
  "q_dgseen",                0,    500, "Network: seen <=14d",
  "q_nrpt",                  1,     95, "Partners /6mo (cap 95)",
  "q_svcevfr",               0,     99, "Coerced-sex count (cap 99)",
  "q_svrevfr",               0,     99, "Forced-sex count (cap 99)",
  "lab_cd4_count",           0,   2000, "CD4 cells/mm3 (phys. plausible)",
  "lab_hiv_viral_load",      0,  1e8,   "Viral load copies/mL",
  "pay_amount",              5,      5, "Fixed incentive = 5",
  "coupon_issued_count",     0,    100, "Coupons issued"
)

range_scan <- pmap_dfr(bounds, function(var, lower, upper, note) {
  x <- gv(wn, var)
  tibble(Variable = var,
         `Non-missing` = sum(!is.na(x)),
         `Below lower` = sum(is_true(x < lower)),
         `Above upper` = sum(is_true(x > upper)),
         Bound = sprintf("[%s, %s]", lower, upper),
         Note = note)
}) %>%
  filter(`Below lower` + `Above upper` > 0)

# lab_cd4_count and pay_amount have dedicated checks below; keep them in the
# table for context but do not also raise a (duplicate) summary issue here.
dedicated <- c("lab_cd4_count", "pay_amount")
if (nrow(range_scan) > 0) {
  for (i in seq_len(nrow(range_scan))) {
    if (range_scan$Variable[i] %in% dedicated) next
    add_issue("Range", "Warning",
              sprintf("%s: %d value(s) outside %s.",
                      range_scan$Variable[i], range_scan$`Below lower`[i] + range_scan$`Above upper`[i],
                      range_scan$Bound[i]))
  }
}
kable_if(range_scan, caption = "Numeric values outside near-impossible bounds")
```

## Suspected ceilings and sentinels

```{r ceilings}
ceil_check <- tribble(~var, ~cap, ~label,
  "q_nrpt", 95, "partners in 6 months",
  "q_svcevfr", 99, "lifetime coercion events",
  "q_svrevfr", 99, "lifetime forced-sex events")

ceil_tab <- pmap_dfr(ceil_check, function(var, cap, label) {
  x <- gv(wn, var)
  n <- sum(!is.na(x)); at_cap <- sum(is_true(x == cap))
  tibble(Variable = var, Cap = cap, `At cap` = at_cap,
         `% at cap` = ifelse(n > 0, round(100 * at_cap / n, 1), NA_real_),
         Meaning = label)
}) %>%
  filter(`At cap` > 0)

if (nrow(ceil_tab) > 0) {
  add_issue("Sentinel/ceiling", "Warning",
            paste0("Pile-up at validation ceiling for: ",
                   paste(ceil_tab$Variable, collapse = ", "),
                   " (suspected caps or sentinels)."))
}
kable_if(ceil_tab, caption = "Counts sitting exactly on a validation ceiling")
```

## Count-variable distribution shape

```{r count-shape}
# Exposure counts should be right-skewed (most respondents low, a few high).
# A median sitting mid-range indicates a flat/uniform shape -- a sign of
# fabricated or sentinel-laden data rather than real reporting.
count_vars <- tribble(~var, ~label,
  "q_nrpt",    "Male anal-sex partners, 6 mo",
  "q_svcevfr", "Lifetime coercion events",
  "q_svrevfr", "Lifetime forced-sex events")

shape_tab <- pmap_dfr(count_vars, function(var, label) {
  x <- gv(wn, var); x <- x[!is.na(x)]
  if (length(x) == 0) return(tibble())
  rng <- range(x); med <- median(x)
  rel <- if (diff(rng) > 0) (med - rng[1]) / diff(rng) else NA_real_
  tibble(Variable = var, Label = label, Min = rng[1], Median = med, Max = rng[2],
         `Median position` = round(rel, 2),
         centered = !is.na(rel) & rel >= 0.35 & rel <= 0.65)
})

flagged_shape <- shape_tab %>% filter(centered)
if (nrow(flagged_shape) > 0) {
  for (i in seq_len(nrow(flagged_shape))) {
    add_issue("Distribution", "Warning",
      sprintf("%s: median %g sits mid-range [%g, %g] (position %.2f of range); count data should be right-skewed, not centered.",
              flagged_shape$Variable[i], flagged_shape$Median[i],
              flagged_shape$Min[i], flagged_shape$Max[i], flagged_shape$`Median position`[i]))
  }
}
kable_if(shape_tab %>% select(-centered),
         caption = "Count-variable distribution shape (median position within observed range)")
```

```{r count-shape-hist, fig.cap="Distributions of count variables flagged as non-skewed"}
if (exists("flagged_shape") && nrow(flagged_shape) > 0) {
  long_counts <- dl %>%
    filter(variable %in% flagged_shape$Variable, !is.na(numeric_value)) %>%
    left_join(count_vars, by = c("variable" = "var"))
  ggplot(long_counts, aes(numeric_value)) +
    geom_histogram(bins = 25, fill = "#756bb1", color = "white") +
    facet_wrap(~ label, scales = "free") +
    theme_minimal() +
    labs(x = "Reported count", y = "Records",
         subtitle = "A flat or centered shape is implausible for exposure counts")
}
```

## Age structure

```{r age-hist, fig.cap="Age distribution (q_tutage)"}
age <- gv(wn, "q_tutage")
age_med <- median(age, na.rm = TRUE)
if (age_med > 45) {
  add_issue("Distribution", "Warning",
            sprintf("Median age = %.0f with a near-uniform spread; not a credible age structure for an MSM sample.", age_med))
}
ggplot(tibble(age = age), aes(age)) +
  geom_histogram(binwidth = 5, fill = "#2c7fb8", color = "white") +
  geom_vline(xintercept = age_med, linetype = "dashed", color = "red") +
  annotate("text", x = age_med, y = Inf, vjust = 2, color = "red",
           label = sprintf("median %.0f", age_med)) +
  theme_minimal() +
  labs(x = "Age (years)", y = "Records",
       subtitle = "A flat 15-89 spread with a high median is atypical for this population")
```

## CD4 plausibility

```{r cd4, fig.cap="CD4 count distribution (lab_cd4_count), HIV-positive only"}
cd4 <- gv(wn, "lab_cd4_count")
cd4 <- cd4[!is.na(cd4)]
n_hi <- sum(cd4 > 2000)
n_frac <- sum(cd4 != round(cd4))
if (length(cd4) > 0) {
  if (n_hi > 0) add_issue("Range", "Critical",
    sprintf("lab_cd4_count: %d value(s) > 2000 cells/mm3 (physiologically implausible).", n_hi))
  if (n_frac > 0) add_issue("Data type", "Warning",
    sprintf("lab_cd4_count: %d non-integer value(s); CD4 is reported as whole cells.", n_frac))
  ggplot(tibble(cd4 = cd4), aes(cd4)) +
    geom_histogram(bins = 25, fill = "#d95f0e", color = "white") +
    geom_vline(xintercept = 2000, linetype = "dashed", color = "red") +
    annotate("text", x = 2000, y = Inf, vjust = 2, hjust = -0.05, color = "red",
             label = "plausible ceiling ~2000") +
    theme_minimal() +
    labs(x = "CD4 (cells/mm3)", y = "Records")
} else cat("No CD4 values present.\n")
```

# Categorical frequency checks

```{r cat-checks, results='asis'}
# Birth sex in an MSM survey
sex <- gv(wt, "q_tutsex")
n_fem <- sum(sex == "Female", na.rm = TRUE)
if (n_fem > 0) {
  add_issue("Eligibility", "Warning",
            sprintf("%d record(s) report birth sex Female (%.1f%%) in an MSM survey.",
                    n_fem, 100 * n_fem / sum(!is.na(sex))))
  cat(sprintf("- **Birth sex:** %d of %d report Female (%.1f%%) — unexpected for an MSM sample.\n\n",
              n_fem, sum(!is.na(sex)), 100 * n_fem / sum(!is.na(sex))))
}

# Never had sex yet completed
eltgsx <- gv(wt, "q_eltgsx")
n_neversex <- sum(eltgsx == "No", na.rm = TRUE)
if (n_neversex > 0) {
  add_issue("Eligibility", "Warning",
            sprintf("%d record(s) report never having had sex yet completed the interview.", n_neversex))
  cat(sprintf("- **Sexual history:** %d report never having had sex (q_eltgsx = No) but completed the survey.\n\n", n_neversex))
}

# Last anal sex with a man > 6 months / don't know among completers
elm <- gv(wt, "q_elmssxt")
n_oob <- sum(elm %in% c("More than 6 months ago", "Don't know"))
if (n_oob > 0) {
  cat(sprintf("- **Recency:** %d completers last had anal sex with a man more than 6 months ago or don't know — outside the typical recent-behaviour window.\n\n", n_oob))
}
```

## Categorical variability scan

```{r flatness}
# Flag the opposite of the usual skew problem: categoricals that are implausibly
# flat (max/min category count near 1) across many variables -> low real variability.
cat_vars <- dl %>%
  filter(!is.na(text_value), grepl("^q_", variable)) %>%
  count(variable, text_value) %>%
  group_by(variable) %>%
  summarise(levels = n(), max_n = max(n), min_n = min(n), total = sum(n), .groups = "drop") %>%
  filter(levels >= 3, total >= 100) %>%
  mutate(flatness = round(max_n / min_n, 2)) %>%
  arrange(flatness)

flat_share <- mean(cat_vars$flatness < 1.6)
if (flat_share > 0.5) {
  add_issue("Distribution", "Info",
            sprintf("%.0f%% of multi-level questions are near-uniform (max/min category < 1.6), consistent with low-variability/synthetic data.",
                    100 * flat_share))
}
knitr::kable(head(cat_vars, 10),
             caption = "Flattest categorical distributions (max/min category count)",
             col.names = c("Variable", "Levels", "Max n", "Min n", "Total", "Max/Min"))
```

# Bivariate consistency checks

## Network-size nesting

```{r nesting}
dgb <- gv(wn, "q_dgbehav"); dga <- gv(wn, "q_dgage"); dgs <- gv(wn, "q_dgseen")
v_sa <- sum(is_true(dgs > dga))   # seen > aged-18+
v_ab <- sum(is_true(dga > dgb))   # aged-18+ > total known
nest <- tibble(
  Constraint = c("q_dgseen <= q_dgage", "q_dgage <= q_dgbehav"),
  Violations = c(v_sa, v_ab)
) %>% filter(Violations > 0)
if (nrow(nest) > 0) {
  add_issue("Consistency", "Critical",
            sprintf("Network nesting violated: %d record(s) with dgseen>dgage, %d with dgage>dgbehav.", v_sa, v_ab))
}
kable_if(nest, caption = "Network-size nesting violations (should be zero)")
```

## HIV rapid vs confirmatory lab

```{r rapid-lab}
rapid <- tolower(gv(wt, "rapid_hivrapid_result"))
conf  <- gv(wt, "lab_hiv_confirmatory_test")
ct <- tibble(rapid = rapid, conf = conf) %>%
  filter(rapid == "positive", !is.na(conf)) %>%
  count(conf) %>% rename(`Confirmatory result` = conf, n = n)
n_discord <- ct %>% filter(`Confirmatory result` == "Negative") %>% pull(n) %>% sum()
if (n_discord > 0) {
  add_issue("Consistency", "Critical",
            sprintf("%d rapid-positive record(s) returned a confirmatory NEGATIVE lab result.", n_discord))
}
knitr::kable(ct, caption = "Confirmatory lab result among rapid-positive records")
```

## Self-reported HIV status vs rapid result

```{r selfreport-hiv, results='asis'}
self_pos <- is_true(gv(wt, "q_tsres") == "Positive") | is_true(gv(wt, "q_tsprovpos") == "Yes")
art_any  <- is_true(gv(wt, "q_artnow") == "Yes") | is_true(gv(wt, "q_artevr") == "Yes")
rp_pos   <- is_true(rapid == "positive")

n_self_pos      <- sum(self_pos)
n_self_pos_rneg <- sum(self_pos & !rp_pos)
n_art           <- sum(art_any)
n_art_rneg      <- sum(art_any & !rp_pos)
n_rapid_pos     <- sum(rp_pos)

if (n_self_pos_rneg > 0)
  add_issue("Consistency", "Warning",
            sprintf("%d record(s) report a positive HIV status (prior test or provider) but are rapid-negative.", n_self_pos_rneg))
if (n_art_rneg > 0)
  add_issue("Consistency", "Warning",
            sprintf("%d record(s) report HIV ARV treatment but are rapid-negative.", n_art_rneg))

cat(sprintf("- Rapid-positive: **%d**. Self-reported HIV-positive: **%d** (of which **%d** rapid-negative).\n", n_rapid_pos, n_self_pos, n_self_pos_rneg))
cat(sprintf("- Report HIV treatment (ever/now ARVs): **%d** (of which **%d** rapid-negative).\n", n_art, n_art_rneg))
```

## Self-reported viral suppression vs lab viral load

```{r vl-consistency}
self_supp <- is_true(gv(wn, "q_csvlrs1_0") == 1) | is_true(gv(wn, "q_csvlrs1_1") == 1)  # Suppressed/Undetectable
lab_vl <- gv(wn, "lab_hiv_viral_load")
lab_unsupp <- is_true(lab_vl >= 1000)
n_conflict <- sum(self_supp & lab_unsupp)
vl_tab <- tibble(
  Group = c("Self-report suppressed/undetectable",
            "  ...with a lab viral load >= 1000 copies/mL"),
  n = c(sum(self_supp), n_conflict))
if (n_conflict > 0) {
  add_issue("Consistency", "Warning",
            sprintf("%d record(s) self-report viral suppression but have a lab viral load >= 1000 copies/mL.", n_conflict))
}
knitr::kable(vl_tab, caption = "Self-reported suppression vs measured viral load")
```

## Skip-logic integrity

```{r skip-logic}
# Lab tests should exist only for rapid-positive records.
lab_present <- !is.na(gv(wn, "lab_cd4_count")) | !is.na(gv(wn, "lab_hiv_viral_load")) |
               !is.na(gv(wt, "lab_hiv_confirmatory_test"))
lab_when_neg <- sum(lab_present & !rp_pos)
lab_missing_when_pos <- sum(!lab_present & rp_pos)

# elmssxt asked only when eltgsx == Yes.
elm_present <- !is.na(gv(wt, "q_elmssxt"))
elm_when_notyes <- sum(elm_present & !is_true(gv(wt, "q_eltgsx") == "Yes"))
elm_missing_when_yes <- sum(!elm_present & is_true(gv(wt, "q_eltgsx") == "Yes"))

skip_tab <- tribble(~Rule, ~Violations,
  "Lab test present but rapid not positive", lab_when_neg,
  "Rapid positive but lab test missing",     lab_missing_when_pos,
  "q_elmssxt answered but q_eltgsx != Yes",  elm_when_notyes,
  "q_eltgsx == Yes but q_elmssxt missing",   elm_missing_when_yes) %>%
  filter(Violations > 0)

if (nrow(skip_tab) > 0) {
  add_issue("Skip logic", "Warning",
            paste0("Skip-pattern violations: ",
                   paste(sprintf("%s (%d)", skip_tab$Rule, skip_tab$Violations), collapse = "; "), "."))
}
kable_if(skip_tab, caption = "Skip-logic violations (should be zero)")
```

# Incentive and timing checks

```{r pay-time, results='asis'}
pay <- gv(wn, "pay_amount")
n_payoff <- sum(is_true(pay != 5))
if (n_payoff > 0) {
  add_issue("Range", "Info",
            sprintf("%d record(s) have a payment amount other than the fixed 5.", n_payoff))
  cat(sprintf("- **Incentive:** %d record(s) paid an amount other than 5 (max observed %s).\n\n",
              n_payoff, max(pay, na.rm = TRUE)))
}

start <- suppressWarnings(ymd_hms(gv(wt, "meta_started_at"), quiet = TRUE))
end   <- suppressWarnings(ymd_hms(gv(wt, "meta_completed_at"), quiet = TRUE))
dur   <- as.numeric(difftime(end, start, units = "mins"))
n_neg   <- sum(is_true(dur < 0))
n_short <- sum(is_true(dur >= 0 & dur < 2))
n_long  <- sum(is_true(dur > 120))
if (n_neg + n_short + n_long > 0) {
  add_issue("Timing", "Warning",
            sprintf("Interview duration anomalies: %d negative, %d under 2 min, %d over 120 min.",
                    n_neg, n_short, n_long))
  cat(sprintf("- **Duration:** %d negative, %d under 2 minutes, %d over 120 minutes.\n\n", n_neg, n_short, n_long))
}

# Completion-date clustering
dcl <- tibble(d = as.Date(end)) %>% filter(!is.na(d)) %>% count(d) %>% arrange(desc(n))
if (nrow(dcl) > 0 && dcl$n[1] / N > 0.25) {
  cat(sprintf("- **Completion clustering:** %.0f%% of records completed on the top 3 dates (%s).\n\n",
              100 * sum(head(dcl$n, 3)) / N,
              paste(format(head(dcl$d, 3)), collapse = ", ")))
}
```

# Findings summary

```{r summary-table}
sev_order <- c(Critical = 1, Warning = 2, Info = 3)
if (nrow(issues) == 0) {
  cat("No data-quality issues were flagged.\n")
} else {
  issues %>%
    mutate(ord = sev_order[Severity]) %>%
    arrange(ord) %>%
    select(-ord) %>%
    knitr::kable(caption = "Flagged data-quality issues, most severe first")
}
```

---

*Conservative screen: values are flagged only against near-impossible bounds, known
ceilings/sentinels, or contradictions between related fields. Counts are aggregates; no
respondent-level records are listed.*
