---
title: "Data Quality Assessment Report"
author: "SALT Management System"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc-depth: 3
    theme: journal
  pdf:
    toc: true
    documentclass: article
  docx:
    toc: true
execute:
  echo: false
  warning: false
  message: false
---

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
library(lubridate)

# Load data
data_long <- read.csv("data_long.csv")
data_wide <- read.csv("data_wide.csv")

# Set up knitr options
knitr::opts_chunk$set(fig.width = 10, fig.height = 6)

# Helper functions
calculate_outliers <- function(x) {
  q1 <- quantile(x, 0.25, na.rm = TRUE)
  q3 <- quantile(x, 0.75, na.rm = TRUE)
  iqr <- q3 - q1
  lower <- q1 - 1.5 * iqr
  upper <- q3 + 1.5 * iqr
  sum(x < lower | x > upper, na.rm = TRUE)
}
```

# Executive Summary

This report provides a comprehensive assessment of data quality for the SALT survey system, identifying potential issues and providing recommendations for improvement.

## Quality Metrics Overview

```{r}
# Calculate overall quality metrics
total_records <- nrow(data_wide)
total_variables <- length(unique(data_long$variable))
total_responses <- nrow(data_long)

# Missing data
missing_count <- sum(is.na(data_long$numeric_value) & is.na(data_long$text_value))
missing_pct <- round((missing_count / total_responses) * 100, 2)

# Duplicates
duplicate_participants <- sum(duplicated(data_wide$participant_id))

# Completeness
avg_completeness <- data_long %>%
  group_by(survey_id) %>%
  summarise(
    total = n(),
    complete = sum(!is.na(numeric_value) | !is.na(text_value))
  ) %>%
  summarise(avg = mean(complete / total * 100)) %>%
  pull(avg) %>%
  round(2)

quality_overview <- data.frame(
  Metric = c(
    "Total Records",
    "Total Variables",
    "Total Data Points",
    "Missing Data %",
    "Average Completeness %",
    "Duplicate Participants"
  ),
  Value = c(
    format(total_records, big.mark = ","),
    total_variables,
    format(total_responses, big.mark = ","),
    paste0(missing_pct, "%"),
    paste0(avg_completeness, "%"),
    duplicate_participants
  ),
  Status = c(
    "Info",
    "Info",
    "Info",
    ifelse(missing_pct < 5, "Good", ifelse(missing_pct < 10, "Warning", "Critical")),
    ifelse(avg_completeness > 90, "Good", ifelse(avg_completeness > 80, "Warning", "Critical")),
    ifelse(duplicate_participants == 0, "Good", "Critical")
  )
)

kable(quality_overview, caption = "Data Quality Overview")
```

# Completeness Analysis

## Missing Data by Variable

```{r}
missing_summary <- data_long %>%
  group_by(variable) %>%
  summarise(
    total = n(),
    missing = sum(is.na(numeric_value) & is.na(text_value)),
    missing_pct = round((missing / total) * 100, 2)
  ) %>%
  arrange(desc(missing_pct))

# Identify problematic variables
problem_vars <- missing_summary %>%
  filter(missing_pct > 20) %>%
  head(20)

if (nrow(problem_vars) > 0) {
  kable(problem_vars,
        caption = "Variables with High Missing Data (>20%)",
        col.names = c("Variable", "Total Records", "Missing", "Missing %"))
} else {
  cat("All variables have less than 20% missing data.")
}
```

## Completeness Distribution

```{r, fig.cap="Distribution of Data Completeness Across Variables"}
missing_summary %>%
  ggplot(aes(x = 100 - missing_pct)) +
  geom_histogram(bins = 30, fill = "#3498db", color = "white") +
  geom_vline(xintercept = 80, linetype = "dashed", color = "red", size = 1) +
  annotate("text", x = 75, y = Inf, label = "80% Target", vjust = 2, color = "red") +
  theme_minimal() +
  labs(title = "Distribution of Variable Completeness",
       x = "Completeness (%)",
       y = "Number of Variables") +
  theme(text = element_text(size = 12))
```

## Completeness by Facility

```{r}
facility_completeness <- data_wide %>%
  left_join(
    data_long %>%
      group_by(survey_id) %>%
      summarise(
        completeness = mean(!is.na(numeric_value) | !is.na(text_value)) * 100
      ),
    by = "survey_id"
  ) %>%
  group_by(meta_facility_name) %>%
  summarise(
    n_surveys = n(),
    avg_completeness = round(mean(completeness, na.rm = TRUE), 2),
    min_completeness = round(min(completeness, na.rm = TRUE), 2),
    max_completeness = round(max(completeness, na.rm = TRUE), 2)
  ) %>%
  arrange(avg_completeness)

kable(facility_completeness,
      caption = "Data Completeness by Facility",
      col.names = c("Facility", "Surveys", "Avg Complete %", "Min %", "Max %"))
```

# Data Consistency

## Duplicate Detection

```{r}
# Check for duplicate participant IDs
duplicates <- data_wide %>%
  group_by(participant_id) %>%
  summarise(count = n()) %>%
  filter(count > 1) %>%
  arrange(desc(count))

if (nrow(duplicates) > 0) {
  kable(head(duplicates, 10),
        caption = "Duplicate Participant IDs Found",
        col.names = c("Participant ID", "Occurrences"))
} else {
  cat("✓ No duplicate participant IDs detected")
}
```

## Temporal Consistency

```{r, fig.cap="Survey Completion Times - Detecting Anomalies"}
data_wide %>%
  mutate(
    completion_time = as.POSIXct(meta_completed_at),
    hour = hour(completion_time),
    date = as.Date(completion_time)
  ) %>%
  ggplot(aes(x = hour)) +
  geom_histogram(bins = 24, fill = "#2ecc71", color = "white") +
  scale_x_continuous(breaks = seq(0, 23, 2)) +
  theme_minimal() +
  labs(title = "Distribution of Survey Completion Times",
       subtitle = "Check for unusual patterns (e.g., surveys at 3 AM)",
       x = "Hour of Day",
       y = "Number of Surveys") +
  theme(text = element_text(size = 12))
```

## Value Range Validation

```{r}
# Check numeric variables for outliers
numeric_vars <- data_long %>%
  filter(!is.na(numeric_value)) %>%
  group_by(variable) %>%
  summarise(
    n = n(),
    mean = round(mean(numeric_value, na.rm = TRUE), 2),
    median = round(median(numeric_value, na.rm = TRUE), 2),
    min = min(numeric_value, na.rm = TRUE),
    max = max(numeric_value, na.rm = TRUE),
    outliers = calculate_outliers(numeric_value),
    outlier_pct = round(outliers / n * 100, 2)
  ) %>%
  filter(outlier_pct > 0) %>%
  arrange(desc(outlier_pct))

if (nrow(numeric_vars) > 0) {
  kable(head(numeric_vars, 15),
        caption = "Numeric Variables with Outliers",
        col.names = c("Variable", "N", "Mean", "Median", "Min", "Max", "Outliers", "Outlier %"))
} else {
  cat("✓ No significant outliers detected in numeric variables")
}
```

# Response Patterns

## Response Distribution Analysis

```{r}
# Analyze response patterns for categorical variables
response_patterns <- data_long %>%
  filter(!is.na(text_value)) %>%
  group_by(variable, text_value) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(variable) %>%
  mutate(
    total = sum(count),
    pct = round(count / total * 100, 2)
  ) %>%
  filter(pct > 90) %>%
  arrange(desc(pct))

if (nrow(response_patterns) > 0) {
  kable(head(response_patterns, 15),
        caption = "Variables with Highly Skewed Responses (>90% single value)",
        col.names = c("Variable", "Value", "Count", "Total", "Percentage"))
  cat("\n⚠ These variables may have data entry issues or low variability\n")
} else {
  cat("✓ No highly skewed response patterns detected")
}
```

## Sequential Pattern Detection

```{r, fig.cap="Survey ID Sequence Check"}
# Check for gaps in survey IDs
survey_ids <- sort(as.numeric(data_wide$survey_id))
expected_ids <- seq(min(survey_ids), max(survey_ids))
missing_ids <- setdiff(expected_ids, survey_ids)

if (length(missing_ids) > 0) {
  cat(paste("⚠", length(missing_ids), "gaps found in survey ID sequence\n"))
  if (length(missing_ids) <= 10) {
    cat(paste("Missing IDs:", paste(head(missing_ids, 10), collapse = ", "), "\n"))
  }
} else {
  cat("✓ Survey ID sequence is complete with no gaps")
}

# Visualize survey ID progression
data.frame(
  index = seq_along(survey_ids),
  survey_id = survey_ids
) %>%
  ggplot(aes(x = index, y = survey_id)) +
  geom_line(color = "#3498db") +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.2) +
  theme_minimal() +
  labs(title = "Survey ID Progression",
       subtitle = "Linear progression indicates consistent data collection",
       x = "Survey Index",
       y = "Survey ID") +
  theme(text = element_text(size = 12))
```

# Data Entry Quality

## Entry Time Analysis

```{r}
# Analyze survey completion duration if available
if ("meta_start_time" %in% names(data_wide) && "meta_completed_at" %in% names(data_wide)) {
  data_wide %>%
    mutate(
      start = as.POSIXct(meta_start_time),
      end = as.POSIXct(meta_completed_at),
      duration_mins = as.numeric(difftime(end, start, units = "mins"))
    ) %>%
    filter(duration_mins > 0 & duration_mins < 180) %>%
    ggplot(aes(x = duration_mins)) +
    geom_histogram(bins = 30, fill = "#e74c3c", color = "white") +
    geom_vline(xintercept = c(5, 60), linetype = "dashed", color = "blue") +
    theme_minimal() +
    labs(title = "Survey Completion Duration",
         subtitle = "Very short (<5 min) or very long (>60 min) durations may indicate issues",
         x = "Duration (minutes)",
         y = "Number of Surveys") +
    theme(text = element_text(size = 12))
}
```

## Facility Performance Scores

```{r}
# Calculate quality scores by facility
facility_scores <- data_wide %>%
  group_by(meta_facility_name) %>%
  summarise(
    n_surveys = n(),
    duplicate_rate = mean(duplicated(participant_id)),
    .groups = "drop"
  ) %>%
  left_join(
    facility_completeness %>% select(meta_facility_name, avg_completeness),
    by = "meta_facility_name"
  ) %>%
  mutate(
    quality_score = round(
      (avg_completeness / 100) * 0.7 +  # 70% weight on completeness
      (1 - duplicate_rate) * 0.3,        # 30% weight on no duplicates
      3
    ) * 100
  ) %>%
  arrange(desc(quality_score))

kable(facility_scores,
      caption = "Facility Data Quality Scores",
      col.names = c("Facility", "Surveys", "Duplicate Rate", "Completeness %", "Quality Score"))
```

# Recommendations

## Critical Issues

```{r}
critical_issues <- character()

if (duplicate_participants > 0) {
  critical_issues <- c(critical_issues,
    paste("•", duplicate_participants, "duplicate participant IDs require immediate investigation"))
}

if (missing_pct > 10) {
  critical_issues <- c(critical_issues,
    paste("• Overall missing data rate of", missing_pct, "% exceeds acceptable threshold"))
}

low_quality_facilities <- facility_scores %>% filter(quality_score < 70)
if (nrow(low_quality_facilities) > 0) {
  critical_issues <- c(critical_issues,
    paste("•", nrow(low_quality_facilities), "facilities have quality scores below 70%"))
}

if (length(critical_issues) > 0) {
  cat("**CRITICAL ISSUES REQUIRING IMMEDIATE ATTENTION:**\n")
  cat(paste(critical_issues, collapse = "\n"))
} else {
  cat("✓ No critical data quality issues detected")
}
```

## Improvement Opportunities

```{r}
recommendations <- character()

# Check for variables with high missing rates
high_missing <- missing_summary %>% filter(missing_pct > 10)
if (nrow(high_missing) > 0) {
  recommendations <- c(recommendations,
    paste("• Review and improve data collection for", nrow(high_missing), "variables with >10% missing data"))
}

# Check for facilities needing support
low_performing <- facility_scores %>% filter(quality_score < 80)
if (nrow(low_performing) > 0) {
  recommendations <- c(recommendations,
    paste("• Provide additional training to", nrow(low_performing), "facilities with quality scores <80%"))
}

# Check for outliers
if (exists("numeric_vars") && nrow(numeric_vars) > 0) {
  recommendations <- c(recommendations,
    paste("• Implement validation rules for", nrow(numeric_vars), "numeric variables with outliers"))
}

if (length(recommendations) > 0) {
  cat("\n**RECOMMENDED IMPROVEMENTS:**\n")
  cat(paste(recommendations, collapse = "\n"))
} else {
  cat("\n✓ Data quality is excellent - maintain current procedures")
}
```

## Quality Trends

```{r, fig.cap="Data Quality Trend Over Time"}
# Calculate quality metrics over time
quality_trend <- data_wide %>%
  mutate(
    date = as.Date(meta_completed_at),
    week = floor_date(date, "week")
  ) %>%
  left_join(
    data_long %>%
      group_by(survey_id) %>%
      summarise(
        completeness = mean(!is.na(numeric_value) | !is.na(text_value)) * 100,
        .groups = "drop"
      ),
    by = "survey_id"
  ) %>%
  group_by(week) %>%
  summarise(
    surveys = n(),
    avg_completeness = mean(completeness, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(!is.na(week))

if (nrow(quality_trend) > 1) {
  ggplot(quality_trend, aes(x = week, y = avg_completeness)) +
    geom_line(color = "#3498db", size = 1) +
    geom_point(color = "#2c3e50", size = 2) +
    geom_smooth(method = "loess", se = TRUE, alpha = 0.2) +
    geom_hline(yintercept = 90, linetype = "dashed", color = "green") +
    scale_y_continuous(limits = c(0, 100)) +
    theme_minimal() +
    labs(title = "Data Quality Trend",
         subtitle = "Weekly average completeness over time",
         x = "Week",
         y = "Completeness (%)") +
    theme(text = element_text(size = 12))
}
```

---

*Report generated on `r format(Sys.Date(), "%B %d, %Y")` at `r format(Sys.time(), "%I:%M %p")`*
*Quality assessment based on `r format(total_records, big.mark = ",")` records*