---
title: "Analysis Report — Short MSM Survey (id 2)"
subtitle: "Full analysis: HIV cascade and topic indicators"
author: "SALT report builder"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc-depth: 3
    theme: journal
    embed-resources: true
    format-links: false
execute:
  echo: false
  warning: false
  message: false
---

```{r preamble, include=FALSE}
# ---------------- Embedded SALT analysis functions ----------------
load_or_install <- function(pkg, repos = "https://cloud.r-project.org") {
  if (!requireNamespace(pkg, quietly = TRUE)) install.packages(pkg, repos = repos)
  library(pkg, character.only = TRUE)
}
load_or_install("tidyverse")
load_or_install("RDS")
load_or_install("survey")
load_or_install("DescTools")

tab <- function(...) table(..., useNA = "always")

single_facility <- function(dat, facility_id, outcome_var, pop_size, subset=NULL, nboot=50){
  set.seed(seed)
  sub <- dat
  if(is.null(subset))
    subset <- rep(TRUE, nrow(dat))
  sub$subset__ <- subset
  sub$subset__[is.na(sub$subset__)] <- FALSE
  sub <- sub |> filter(meta_facility_id == facility_id)

  rds_df <- as.rds.data.frame(sub,
                              id = id_var_name,
                              recruiter.id = recruiter_id_name,
                              network.size = degree_name,
                              time = time_var_name,
                              population.size = pop_size)

  wts <- RDS::compute.weights(rds_df, outcome.variable="subset__", estimate_method, N = pop_size)
  v <- !rds_df$subset__ | is.na(rds_df[[outcome_var]])

  domain_probs <- sum(rds_df$subset__* wts, na.rm=TRUE) / sum(wts[!is.na(rds_df$subset__)])
  if(nrow(rds_df) == 0 || all(v, na.rm=TRUE)){
    if(is.factor(rds_df[[outcome_var]])){
      nlev <- length(levels(rds_df[[outcome_var]]))
      n <- rep(0,nlev)
    }else{
      nlev <- 1
      n <- 0
    }
    est <- rep(as.numeric(NA), nlev)
    names(est) <- levels(rds_df[[outcome_var]])
    interval <- 1
    attr(interval, "bsresult") <- list(se_estimate = rep(as.numeric(NA), nlev))
    boot_est <- list(
      estimate = est,
      interval=interval,
      n=n,
      domain_probs = domain_probs*0
    )
    return(boot_est)
  }
  rds_df$outcome__ <- rds_df[[outcome_var]]
  rds_df$outcome__[!rds_df$subset__] <- NA
  nn <- sum(!is.na(rds_df$outcome__))
  tr <- try(boot_est <- RDS.bootstrap.intervals(rds.data = rds_df,
                                                weight.type = estimate_method,
                                                uncertainty = bootstrap_method,
                                                outcome.variable = "outcome__",
                                                N = pop_size,
                                                number.of.bootstrap.samples = nboot))
  if(inherits(tr,"try-error")){
    warning(tr)
    boot_est <- NULL
  }
  if(is.null(boot_est)){
    wts <- RDS::compute.weights(rds_df, outcome.variable = "outcome__", weight.type = estimate_method, N = pop_size)
    sub$weights__ <- wts
    df <- sub[sub$subset__ & !is.na(sub[[outcome_var]]),]
    df$outcome_var__ <- df[[outcome_var]]
    if(nrow(df) == 1){
      if(is.factor(df$outcome_var__)){
        est <- table(df[[outcome_var]]) |> as.numeric()
        names(est) <- levels(df[[outcome_var]])
        se <- sqrt(.5*.5)
      }else{
        est <- df[[outcome_var]]
        se <- NA
      }
    }else{
      des <- svydesign(ids=~1, data=df, weights = ~weights__)
      res <- svymean(~outcome_var__,design = des, na.rm=TRUE)
      est <- as.vector(res)
      names(est) <- levels(df[[outcome_var]])
      se <- sqrt(diag(attr(res,"var")))
    }
    if(is.factor(df$outcome_var__)){
      for(i in 1:length(est)){
        if(est[i] < .00001){
          bi <- BinomCI(0, nrow(df), method ="clopper-pearson")
          se_cp <- (bi[3] - bi[2]) / (1.96)
          se[i] <- se_cp
        }
        if(est[i] > .999999){
          bi <- BinomCI(nrow(df), nrow(df), method ="clopper-pearson")
          se_cp <- (bi[3] - bi[2]) / (1.96)
          se[i] <- se_cp
        }
      }
      nn <- table(df$outcome_var__) |> as.vector()
    }else{
      nn <- length(na.omit(df$outcome_var__))
    }
    interval <- 1
    attr(interval, "bsresult") <- list(se_estimate = se)
    boot_est <- list(
      estimate = est,
      interval=interval,
      n=nn,
      domain_probs = domain_probs
    )
  }else{
    if(is.factor(rds_df$outcome__))
      boot_est$n <- matrix(boot_est$interval, ncol = 6, byrow = FALSE)[,6]
    else
      boot_est$n <- nn
    boot_est$domain_probs <- domain_probs
  }
  if(!is.factor(dat[[outcome_var]])){
    v <- sub[sub$subset__,outcome_var]
    raw_mean <- mean(v, na.rm=TRUE)
  }else{
    raw_mean <- as.numeric(NA)
  }
  results <- data.frame(facility_id=facility_id,estimate_name=names(boot_est$estimate),estimate=boot_est$estimate,
                    se=attr(boot_est$interval,"bsresult")$se_estimate,
                    n=boot_est$n,domain_probs=boot_est$domain_probs)
  res <- data.frame(
    estimate_name = results$estimate_name,
    estimate = results$estimate,
    se = results$se,
    lower = results$estimate - 1.96 * results$se,
    upper = results$estimate + 1.96 * results$se,
    raw_mean = raw_mean,
    n = results$n
  )
  res
}

all_facilities <- function(dat, outcome_var, size_by_facility, stratify_by=NULL, subset=NULL, nboot=50){
  if(!is.null(stratify_by)){
    results <- list()
    if(!is.factor(dat[[stratify_by]])){
      dat[[stratify_by]] <- as.factor(dat[[stratify_by]])
    }
    svar <- dat[[stratify_by]]
    levs <- levels(svar)
    if(is.null(subset))
      subset <- rep(TRUE, nrow(dat))
    for(l in levs){
      sub <- subset & svar == l
      res <- all_facilities(dat=dat, outcome_var=outcome_var,
                            size_by_facility=size_by_facility, subset=sub, nboot=nboot)
      res$stratify_value <- l
      results[[length(results) + 1]] <- res
    }
    results <- bind_rows(results)
    results$stratify_variable <- stratify_by
    return(results)
  }

  if(is.character(dat[[outcome_var]])){
    dat[[outcome_var]] <- as.factor(dat[[outcome_var]])
  }
  results <- list()
  facilities <- unique(dat$meta_facility_id)
  for(f in facilities){
    pop_size <- size_by_facility$size[size_by_facility$facility_id == f]
    r <- single_facility(dat, facility_id=f, outcome_var=outcome_var, pop_size, subset=subset, nboot=nboot)
    r$facility_id <- f
    results[[as.character(f)]] <- r
  }
  results <- bind_rows(results)
  results$outcome <- outcome_var

  totals <- results |>
    left_join(size_by_facility, by = "facility_id") |>
    group_by(estimate_name) |>
    mutate(p = representation_proportion / sum(representation_proportion)) |>
    summarise(
      estimate = sum(estimate * p),
      se = sqrt( sum(se^2 * p^2) ),
      raw_mean = sum(raw_mean[n>0] * n[n>0]/ sum(n[n>0])) ,
      n = sum(n),
      outcome=outcome[1],
    ) |>
    mutate(
      lower = estimate - 1.96 * se,
      upper = estimate + 1.96 * se,
      facility_id = -1
    )
  results <- bind_rows(results,totals)
  results
}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

dat <- read.csv("data_wide.csv",         stringsAsFactors = FALSE, check.names = FALSE)
dwn <- read.csv("data_wide_numeric.csv", stringsAsFactors = FALSE, check.names = FALSE)

idc  <- intersect(c("participant_id", "meta_participant_id"), names(dat))[1]
idc2 <- intersect(c("participant_id", "meta_participant_id"), names(dwn))[1]
if (!("participant_id" %in% names(dat))) dat$participant_id <- dat[[idc]]

get_txt <- function(v) if (v %in% names(dat)) as.character(dat[[v]]) else rep(NA_character_, nrow(dat))
get_num <- function(v) {
  if (!(v %in% names(dwn))) return(rep(NA_real_, nrow(dat)))
  idx <- match(dat$participant_id, dwn[[idc2]])
  suppressWarnings(as.numeric(dwn[[v]][idx]))
}
eqT <- function(x, val) !is.na(x) & x == val

# Interview start time, parsed for the RDS time= argument (handles ISO8601)
dat$meta_started_at <- lubridate::as_datetime(dat$meta_started_at, tz = "UTC")

# ---- Recruiter id and adjusted, capped network size ------------------------
dat$recruiter_id__ <- rid.from.coupons(
  dat, subject.coupon = "coupon_referral_used",
  coupon.variables = paste0("coupon_issued_", 1:3),
  subject.id = "participant_id")

is_seed    <- !(dat$recruiter_id__ %in% dat$participant_id)
n_recruits <- as.integer(table(factor(dat$recruiter_id__[!is_seed], levels = dat$participant_id)))
n_recruits[is.na(n_recruits)] <- 0L

deg <- get_num("q_dgseen")
adj <- pmax(deg, n_recruits)
adj <- adj + ifelse(is_seed, 0L, 1L)
adj[adj == 0] <- 1
adj[adj < 0]  <- NA
facn <- get_txt("meta_facility_name")
medf <- tapply(deg, facn, median, na.rm = TRUE)
na_idx <- which(is.na(adj)); adj[na_idx] <- as.numeric(medf[facn[na_idx]])
adj[is.na(adj)] <- median(deg, na.rm = TRUE)
n_capped <- sum(adj > 100, na.rm = TRUE)
adj[adj > 100] <- 100                      # ceiling at 100 (per request)
dat$degree__ <- adj

if ("meta_facility_id" %in% names(dat)) {
  dat$meta_facility_id <- suppressWarnings(as.integer(dat$meta_facility_id))
} else {
  dat$meta_facility_id <- suppressWarnings(as.integer(get_num("meta_facility_id")))
}

# ---- Cascade variables (factors) -------------------------------------------
rapid_pos <- tolower(get_txt("rapid_hivrapid_result")) == "positive"
conf      <- get_num("lab_hiv_confirmatory_test")           # 0=Pos 1=Neg 2=Indet
hiv_pos   <- ifelse(!is.na(conf) & conf %in% c(0, 1), conf == 0, rapid_pos)
art_bin   <- eqT(get_num("q_artnow"), 0)
diag_bin  <- ifelse(hiv_pos %in% TRUE,
                    eqT(get_num("q_tsres"), 0) | eqT(get_num("q_tsprovpos"), 0) | art_bin, NA)
onart_bin <- ifelse(hiv_pos %in% TRUE, art_bin, NA)
vl        <- get_num("lab_hiv_viral_load")
supp_bin  <- ifelse(hiv_pos %in% TRUE, ifelse(is.na(vl), NA, vl <= 1000), NA)

dat$hiv_f   <- factor(ifelse(is.na(hiv_pos),  NA, ifelse(hiv_pos,  "HIV+",       "HIV-")),        levels = c("HIV-", "HIV+"))
dat$diag_f  <- factor(ifelse(is.na(diag_bin), NA, ifelse(diag_bin, "Diagnosed",  "Not diagnosed")), levels = c("Not diagnosed", "Diagnosed"))
dat$onart_f <- factor(ifelse(is.na(onart_bin),NA, ifelse(onart_bin,"On ART",     "Not on ART")),  levels = c("Not on ART", "On ART"))
dat$supp_f  <- factor(ifelse(is.na(supp_bin), NA, ifelse(supp_bin, "Suppressed", "Not suppressed")), levels = c("Not suppressed", "Suppressed"))

# Age bands (10-year)
age <- get_num("q_tutage")
dat$age_band <- cut(age, breaks = c(15, 25, 35, 45, 55, 65, 75, 90), right = FALSE,
                    labels = c("15-24", "25-34", "35-44", "45-54", "55-64", "65-74", "75+"))

# Subpopulation subsets for topic tables
hivpos_s    <- hiv_pos %in% TRUE
hivneg_s    <- hiv_pos %in% FALSE
drinkers_s  <- get_num("q_alcfreq") %in% c(1, 2, 3, 4)   # not "Never"
injected_s  <- eqT(get_num("q_elinj"), 0)                # ever injected
heardprep_s <- eqT(get_num("q_pprknow"), 0)              # heard of PrEP
onart_s     <- eqT(get_num("q_artnow"), 0)               # currently on ART

# Multi-select indicators -> binary factors (self-reported viral-load result)
ms_cols <- grep("^q_csvlrs1_[0-9]+$", names(dat), value = TRUE)
ms_labels <- c(q_csvlrs1_0 = "Suppressed", q_csvlrs1_1 = "Undetectable",
               q_csvlrs1_2 = "Not suppressed", q_csvlrs1_3 = "Result not given",
               q_csvlrs1_4 = "Don't know", q_csvlrs1_5 = "Prefer not to answer")
for (cc in ms_cols) {
  dat[[cc]] <- factor(ifelse(get_num(cc) == 1, "Selected", "Not selected"),
                      levels = c("Not selected", "Selected"))
}

nb_full <- 25   # bootstrap resamples for the comprehensive run (raise to 50 for a final)

# ---- Globals required by the engine ----------------------------------------
id_var_name       <- "participant_id"
recruiter_id_name <- "recruiter_id__"
degree_name       <- "degree__"
time_var_name     <- "meta_started_at"
estimate_method   <- "HCG"
bootstrap_method  <- "Gile"
seed              <- 100
nboot             <- 1000

# ---- Facility sizes / representation ---------------------------------------
fac_map <- dat |> distinct(meta_facility_id, meta_facility_name) |>
  arrange(meta_facility_id) |> filter(!is.na(meta_facility_id))
fac_n <- as.integer(table(factor(dat$meta_facility_id, levels = fac_map$meta_facility_id)))
rep_map <- c("Demo Facility" = .2, "East Clinic" = .2, "North Clinic" = .3, "South Clinic" = .25)
size_by_facility <- data.frame(
  facility_id = fac_map$meta_facility_id,
  size = 1000 * fac_n,
  representation_proportion = as.numeric(rep_map[fac_map$meta_facility_name])
)
fac_label <- function(fid) if (fid == -1) "Aggregate" else fac_map$meta_facility_name[fac_map$meta_facility_id == fid]
```

# Background

*To be completed once a study protocol is provided.*

# Methods

This analysis treats the survey as a respondent-driven sample. Recruiter links are reconstructed
from the referral and issued coupon codes (`rid.from.coupons`). Each subject's network size is the
reported degree (`q_dgseen`) adjusted to be at least the number of people they recruited, plus one
for non-seeds (the recruiter is assumed not to be counted in the reported degree); zeros are set to
one, negatives to missing, and missing values filled with the facility-median degree. Adjusted
degrees above 100 are capped at 100 (`r n_capped` record(s) affected).

Point estimates use the **HCG** (homophily configuration graph) estimator with finite-population
correction; uncertainty uses the **Gile bootstrap** (`r nboot` resamples for the cascade). When the
bootstrap cannot run — typically very low cell counts — the estimate falls back to a
weighted `survey` design mean, with Clopper-Pearson standard errors at 0%/100% cells. No external
population size estimates were available, so each facility's population is treated as large relative
to its sample (the sample is a small fraction of the population). Aggregate estimates pool facilities
by their supplied representation proportions; their standard errors combine the facility variances
under independence. Topic-indicator tables use the same HCG estimator with `r nb_full` bootstrap
resamples (reduced from the cascade's `r nboot` to keep the comprehensive run tractable), each
computed on the subpopulation stated in its table title.

**HIV cascade definitions.** HIV status uses the rapid test, overridden by the confirmatory lab when
that result is Positive or Negative (an Indeterminate confirmatory falls back to the rapid result;
an indeterminate rapid would be treated as missing). Awareness ("diagnosed", among HIV-positives) is
a self-reported prior positive test, a provider-reported positive, or current ART. On ART is current
ART for treatment (`q_artnow`, excluding PrEP). Viral suppression is a viral load ≤ 1000 copies/mL.
The ARV-drug-test conditions in the cascade specification are omitted because no such assay is
collected in this survey. The conditional cascade chains denominators (aware among positive → on ART
among aware → suppressed among on ART); the unconditional cascade expresses on-ART and suppression as
shares of all HIV-positives.

# Results

```{r run-cascade, include=FALSE}
run_q <- function(outcome_col, pos_level, subset) {
  res <- tryCatch(
    all_facilities(dat = dat, outcome_var = outcome_col, size_by_facility = size_by_facility,
                   subset = subset, nboot = nboot),
    error = function(e) NULL)
  facs <- c(size_by_facility$facility_id, -1)
  do.call(rbind, lapply(facs, function(f) {
    if (is.null(res)) return(data.frame(facility_id = f, n = NA, N = NA, sample_pct = NA,
                                        est_pct = NA, lo = NA, hi = NA))
    rf <- res[res$facility_id == f, ]
    denom <- sum(rf$n, na.rm = TRUE)
    rl <- rf[rf$estimate_name == pos_level, ]
    has <- nrow(rl) == 1 && !is.na(rl$estimate[1])
    data.frame(
      facility_id = f,
      n  = if (nrow(rl)) rl$n[1] else NA,
      N  = denom,
      sample_pct = if (denom > 0 && nrow(rl)) 100 * rl$n[1] / denom else NA,
      est_pct = if (has) 100 * rl$estimate[1] else NA,
      lo = if (has) 100 * rl$lower[1] else NA,
      hi = if (has) 100 * rl$upper[1] else NA)
  }))
}

prevalence   <- run_q("hiv_f",   "HIV+",       NULL)
diag_hivpos  <- run_q("diag_f",  "Diagnosed",  hiv_pos %in% TRUE)
onart_diag   <- run_q("onart_f", "On ART",     diag_bin %in% TRUE)
supp_onart   <- run_q("supp_f",  "Suppressed", onart_bin %in% TRUE)
onart_hivpos <- run_q("onart_f", "On ART",     hiv_pos %in% TRUE)
supp_hivpos  <- run_q("supp_f",  "Suppressed", hiv_pos %in% TRUE)

quantities <- list(
  list(name = "HIV prevalence (of all participants)",        res = prevalence),
  list(name = "Aware of status | HIV+ (95-1)",               res = diag_hivpos),
  list(name = "On ART | aware (95-2, conditional)",          res = onart_diag),
  list(name = "Suppressed | on ART (95-3, conditional)",     res = supp_onart),
  list(name = "On ART | HIV+ (unconditional)",               res = onart_hivpos),
  list(name = "Suppressed | HIV+ (unconditional)",           res = supp_hivpos))

fmt_cell <- function(x, fmt = "%.1f") ifelse(is.na(x), "—", sprintf(fmt, x))
build_fac_table <- function(fid) {
  do.call(rbind, lapply(quantities, function(q) {
    r <- q$res[q$res$facility_id == fid, ]
    data.frame(
      Measure = q$name,
      `n/N` = if (is.na(r$N) || is.na(r$n)) "—" else sprintf("%d/%d", as.integer(r$n), as.integer(r$N)),
      `Sample %` = fmt_cell(r$sample_pct),
      `Pop. % (HCG)` = fmt_cell(r$est_pct),
      `95% CI` = if (is.na(r$lo)) "—" else sprintf("%.1f–%.1f", pmax(0, r$lo), pmin(100, r$hi)),
      check.names = FALSE)
  }))
}
```

## HIV cascade of care (95-95-95)

Each table reports, per facility and pooled, the sample numerator/denominator, the raw sample
percentage, the HCG population estimate, and its 95% bootstrap interval. The conditional cascade
chains denominators; the unconditional rows express on-ART and suppression among all HIV-positives.

```{r cascade-tables, results='asis'}
for (fid in c(size_by_facility$facility_id, -1)) {
  cat(sprintf("\n\n### %s\n\n", fac_label(fid)))
  print(knitr::kable(build_fac_table(fid), row.names = FALSE,
                     caption = sprintf("HIV cascade — %s", fac_label(fid))))
  cat("\n\n")
}
```

## HIV prevalence by age band (aggregate)

The age-stratified table is limited to HIV prevalence; the ART and suppression rungs have too few
HIV-positive cases per age band to estimate meaningfully on this data. 10-year bands are used (none
fall below the collapse threshold for prevalence).

```{r prevalence-age, results='asis'}
res_age <- tryCatch(
  all_facilities(dat = dat, outcome_var = "hiv_f", stratify_by = "age_band",
                 size_by_facility = size_by_facility, nboot = nboot),
  error = function(e) NULL)

if (is.null(res_age)) {
  cat("*Age-stratified estimation was not available.*\n\n")
} else {
  bands <- levels(dat$age_band)
  age_tab <- do.call(rbind, lapply(bands, function(b) {
    rb <- res_age[res_age$facility_id == -1 & res_age$stratify_value == b, ]
    denom <- sum(rb$n, na.rm = TRUE)
    rl <- rb[rb$estimate_name == "HIV+", ]
    has <- nrow(rl) == 1 && !is.na(rl$estimate[1])
    data.frame(
      `Age band` = b,
      `n/N` = if (denom > 0 && nrow(rl)) sprintf("%d/%d", as.integer(rl$n[1]), as.integer(denom)) else "—",
      `Sample %` = if (denom > 0 && nrow(rl)) sprintf("%.1f", 100 * rl$n[1] / denom) else "—",
      `Pop. % (HCG)` = if (has) sprintf("%.1f", 100 * rl$estimate[1]) else "—",
      `95% CI` = if (has) sprintf("%.1f–%.1f", pmax(0, 100 * rl$lower[1]), pmin(100, 100 * rl$upper[1])) else "—",
      check.names = FALSE)
  }))
  print(knitr::kable(age_tab, row.names = FALSE,
                     caption = "Aggregate HIV prevalence by age band (HCG)"))
  cat("\n\n")
}
```

## Topic indicators

Each table summarises one topic at the aggregate level: the sample count and percentage for every
response level, and the HCG population estimate with its 95% interval. The subpopulation each table
is computed on is stated in its title. Multi-select items show one row per option; continuous items
show the sample mean.

```{r topic-engine, include=FALSE}
agg_levels <- function(col, subset = NULL) {
  res <- tryCatch(all_facilities(dat, col, size_by_facility, subset = subset, nboot = nb_full),
                  error = function(e) NULL)
  if (is.null(res)) return(NULL)
  ra <- res[res$facility_id == -1, ]
  denom <- sum(ra$n, na.rm = TRUE)
  data.frame(level = as.character(ra$estimate_name), n = ra$n, denom = denom,
             sample_pct = ifelse(denom > 0, 100 * ra$n / denom, NA),
             est_pct = 100 * ra$estimate, lo = 100 * ra$lower, hi = 100 * ra$upper,
             stringsAsFactors = FALSE)
}
ci_str <- function(lo, hi) ifelse(is.na(lo), "—", sprintf("%.1f–%.1f", pmax(0, lo), pmin(100, hi)))
num_get <- function(col) get_num(col)

build_topic <- function(items, subset) {
  rows <- list()
  add <- function(...) rows[[length(rows) + 1]] <<- data.frame(..., check.names = FALSE)
  for (it in items) {
    sb <- if (!is.null(it$subset)) it$subset else subset
    if (it$type == "num") {
      x <- num_get(it$col); if (!is.null(sb)) x <- x[sb]
      add(Measure = paste0(it$label, " (mean)"), n = sum(!is.na(x)),
          `Sample %/mean` = sprintf("%.1f", mean(x, na.rm = TRUE)),
          `Pop. % (HCG)` = "—", `95% CI` = "—")
    } else if (it$type == "multi") {
      al <- agg_levels(it$col, sb); sel <- if (!is.null(al)) al[al$level == "Selected", ] else NULL
      ok <- !is.null(sel) && nrow(sel) == 1
      add(Measure = it$label, n = if (ok) as.integer(sel$n) else NA,
          `Sample %/mean` = if (ok) sprintf("%.1f", sel$sample_pct) else "—",
          `Pop. % (HCG)` = if (ok && !is.na(sel$est_pct)) sprintf("%.1f", sel$est_pct) else "—",
          `95% CI` = if (ok) ci_str(sel$lo, sel$hi) else "—")
    } else {
      al <- agg_levels(it$col, sb)
      add(Measure = paste0("**", it$label, "**"), n = NA,
          `Sample %/mean` = "", `Pop. % (HCG)` = "", `95% CI` = "")
      if (!is.null(al)) for (i in seq_len(nrow(al))) {
        add(Measure = paste0("  ", al$level[i]), n = as.integer(al$n[i]),
            `Sample %/mean` = if (!is.na(al$sample_pct[i])) sprintf("%.1f", al$sample_pct[i]) else "—",
            `Pop. % (HCG)` = if (!is.na(al$est_pct[i])) sprintf("%.1f", al$est_pct[i]) else "—",
            `95% CI` = ci_str(al$lo[i], al$hi[i]))
      }
    }
  }
  out <- do.call(rbind, rows)
  out$n <- ifelse(is.na(out$n), "", as.character(out$n))
  out
}
```

```{r topics, results='asis'}
ever_tested_s <- eqT(get_num("q_tsever"), 0)
prep_use_s    <- hivneg_s & heardprep_s

topics <- list(
  list(title = "Demographics", sublab = "all participants", subset = NULL, items = list(
    list(col = "q_tutsex",   type = "cat", label = "Sex at birth"),
    list(col = "q_tutage",   type = "num", label = "Age, years"),
    list(col = "q_tutedu",   type = "cat", label = "Highest education"),
    list(col = "q_tutnat",   type = "cat", label = "Nationality"),
    list(col = "q_demarsta", type = "cat", label = "Marital status"))),
  list(title = "Sexual behaviour and violence", sublab = "all participants", subset = NULL, items = list(
    list(col = "q_nrpt",    type = "num", label = "Male anal-sex partners, 6 mo"),
    list(col = "q_cond",    type = "cat", label = "Condom at last anal sex"),
    list(col = "q_trsex",   type = "cat", label = "Ever transactional anal sex"),
    list(col = "q_svcevfr", type = "num", label = "Times coerced into sex (lifetime)"),
    list(col = "q_svrevfr", type = "num", label = "Times forced into sex (lifetime)"))),
  list(title = "Alcohol use", sublab = "all participants (quantity items among drinkers)", subset = NULL, items = list(
    list(col = "q_alcfreq",  type = "cat", label = "Drinking frequency"),
    list(col = "q_alcday",   type = "cat", label = "Drinks per drinking day", subset = drinkers_s),
    list(col = "q_alcbinge", type = "cat", label = "Binge frequency", subset = drinkers_s))),
  list(title = "Drug use", sublab = "all (needle-sharing among those who ever injected)", subset = NULL, items = list(
    list(col = "q_elinj",   type = "cat", label = "Ever injected drugs"),
    list(col = "q_drshare", type = "cat", label = "Ever shared needles", subset = injected_s))),
  list(title = "Mental health", sublab = "all participants", subset = NULL, items = list(
    list(col = "q_phq1", type = "cat", label = "Little interest/pleasure (2 wk)"),
    list(col = "q_phq2", type = "cat", label = "Feeling down/depressed (2 wk)"),
    list(col = "q_mhys", type = "cat", label = "Ever self-harmed"))),
  list(title = "Health care and stigma", sublab = "all participants", subset = NULL, items = list(
    list(col = "q_stgdeny",  type = "cat", label = "Denied care (12 mo)"),
    list(col = "q_hctell",   type = "cat", label = "Comfortable disclosing MSM to provider"),
    list(col = "q_stgavoid", type = "cat", label = "Avoided care for fear of discrimination"),
    list(col = "q_csonline", type = "cat", label = "Accessed HIV/sexual-health info online"),
    list(col = "q_outexp",   type = "cat", label = "Reached by outreach/peer educator"))),
  list(title = "HIV testing history", sublab = "all (test-history items among ever-tested)", subset = NULL, items = list(
    list(col = "q_tsever",     type = "cat", label = "Ever tested for HIV"),
    list(col = "q_tsetwlmth",  type = "cat", label = "Tested in past 12 months", subset = ever_tested_s),
    list(col = "q_tsres",      type = "cat", label = "Last test result", subset = ever_tested_s),
    list(col = "q_tsprovpos",  type = "cat", label = "Provider ever said HIV+", subset = ever_tested_s))),
  list(title = "ART and viral load", sublab = "HIV-positive (treatment-detail items among those on ART)", subset = hivpos_s, items = list(
    list(col = "q_artevr",  type = "cat", label = "Ever taken ARVs"),
    list(col = "q_artnow",  type = "cat", label = "Currently on ARVs"),
    list(col = "q_artprov", type = "cat", label = "Where ARVs obtained", subset = onart_s),
    list(col = "q_artdup",  type = "cat", label = "Collects ARVs from multiple places", subset = onart_s),
    list(col = "q_uuvltest",type = "cat", label = "Viral-load test history", subset = onart_s),
    list(col = "q_csvlrs1_0", type = "multi", label = "Self-report VL: Suppressed", subset = onart_s),
    list(col = "q_csvlrs1_1", type = "multi", label = "Self-report VL: Undetectable", subset = onart_s),
    list(col = "q_csvlrs1_2", type = "multi", label = "Self-report VL: Not suppressed", subset = onart_s),
    list(col = "q_uuknow",  type = "cat", label = "Heard of U=U"),
    list(col = "q_uubel",   type = "cat", label = "Belief in U=U"))),
  list(title = "PrEP", sublab = "HIV-negative (access items among those who have heard of PrEP)", subset = hivneg_s, items = list(
    list(col = "q_pprknow", type = "cat", label = "Heard of PrEP"),
    list(col = "q_proffer", type = "cat", label = "Offered PrEP by provider", subset = prep_use_s),
    list(col = "q_prget",   type = "cat", label = "Knows where to get PrEP", subset = prep_use_s),
    list(col = "q_prtake",  type = "cat", label = "Ever used PrEP", subset = prep_use_s),
    list(col = "q_prwant",  type = "cat", label = "Wants to use PrEP"))),
  list(title = "TB", sublab = "HIV-positive participants", subset = hivpos_s, items = list(
    list(col = "q_tbscrnp", type = "cat", label = "Screened for TB symptoms"),
    list(col = "q_tbdx",    type = "cat", label = "TB diagnosis told"),
    list(col = "q_tbpt",    type = "cat", label = "Given TB-preventive drugs"),
    list(col = "q_tbtx",    type = "cat", label = "Treated for TB"))),
  list(title = "Other STIs and syphilis", sublab = "all participants", subset = NULL, items = list(
    list(col = "q_stdiag",               type = "cat", label = "Told had a non-HIV STI (12 mo)"),
    list(col = "rapid_syprapid_result",  type = "cat", label = "Syphilis rapid test")))
)

for (tp in topics) {
  cat(sprintf("\n\n### %s\n\n", tp$title))
  tb <- tryCatch(build_topic(tp$items, tp$subset), error = function(e) NULL)
  if (is.null(tb)) { cat("*Table could not be produced for this topic.*\n\n"); next }
  print(knitr::kable(tb, row.names = FALSE,
        caption = sprintf("%s — subpopulation: %s (aggregate; %d bootstrap resamples)",
                          tp$title, tp$sublab, nb_full)))
  cat("\n\n")
}
```

# Discussion

*To be completed once the user requests narrative.*

# Conclusion

*To be completed once the user requests narrative.*

# References

*None cited.*
