---
title: "Weekly Activity Report"
subtitle: "Week of `r format(Sys.Date() - 7, '%B %d, %Y')` to `r format(Sys.Date(), '%B %d, %Y')`"
author: "SALT Management System"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc-depth: 2
    theme: flatly
  pdf:
    toc: true
    documentclass: report
  docx:
    toc: true
execute:
  echo: false
  warning: false
  message: false
---

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
library(lubridate)
library(scales)

# Load data
data_long <- read.csv("data_long.csv")
data_wide <- read.csv("data_wide.csv")

# Convert date columns
data_wide$meta_completed_at <- as.POSIXct(data_wide$meta_completed_at)
data_long$meta_completed_at <- as.POSIXct(data_long$meta_completed_at)

# Filter for last 7 days
week_end <- Sys.Date()
week_start <- week_end - 6
week_data <- data_wide %>%
  filter(as.Date(meta_completed_at) >= week_start & as.Date(meta_completed_at) <= week_end)

week_data_long <- data_long %>%
  filter(survey_id %in% week_data$survey_id)

# Set up knitr options
knitr::opts_chunk$set(fig.width = 10, fig.height = 6)
```

# Weekly Summary

## Overview

```{r}
# Calculate key metrics
total_surveys <- nrow(week_data)
total_participants <- length(unique(week_data$participant_id))
active_facilities <- length(unique(week_data$meta_facility_name))

# Compare to previous week
prev_week_data <- data_wide %>%
  filter(as.Date(meta_completed_at) >= (week_start - 7) & as.Date(meta_completed_at) < week_start)

prev_surveys <- nrow(prev_week_data)
change_pct <- ifelse(prev_surveys > 0,
                     round((total_surveys - prev_surveys) / prev_surveys * 100, 1),
                     NA)

summary_stats <- data.frame(
  Metric = c("Surveys Completed", "Unique Participants", "Active Facilities", "Daily Average", "Week-over-Week Change"),
  Value = c(
    total_surveys,
    total_participants,
    active_facilities,
    round(total_surveys / 7, 1),
    ifelse(is.na(change_pct), "N/A", paste0(ifelse(change_pct > 0, "+", ""), change_pct, "%"))
  )
)

kable(summary_stats, caption = "Weekly Performance Metrics")
```

# Daily Activity

## Survey Completions by Day

```{r}
daily_completions <- week_data %>%
  mutate(date = as.Date(meta_completed_at)) %>%
  group_by(date) %>%
  summarise(
    surveys = n(),
    participants = n_distinct(participant_id),
    facilities = n_distinct(meta_facility_name)
  ) %>%
  arrange(date)

# Add day of week
daily_completions$weekday <- weekdays(daily_completions$date, abbreviate = TRUE)

kable(daily_completions,
      caption = "Daily Survey Activity",
      col.names = c("Date", "Surveys", "Participants", "Facilities", "Day"))
```

## Activity Trend

```{r, fig.cap="Daily Survey Completions This Week"}
daily_completions %>%
  ggplot(aes(x = date, y = surveys)) +
  geom_line(color = "#3498db", size = 1.5) +
  geom_point(color = "#2c3e50", size = 3) +
  geom_text(aes(label = surveys), vjust = -1, size = 3.5) +
  scale_x_date(date_breaks = "1 day", date_labels = "%a\n%m/%d") +
  scale_y_continuous(breaks = pretty_breaks()) +
  theme_minimal() +
  labs(title = "Daily Survey Completions",
       subtitle = paste("Week of", format(week_start, "%B %d"), "-", format(week_end, "%B %d, %Y")),
       x = "Date",
       y = "Number of Surveys") +
  theme(text = element_text(size = 12))
```

# Facility Performance

## Activity by Facility

```{r}
facility_activity <- week_data %>%
  group_by(meta_facility_name) %>%
  summarise(
    surveys = n(),
    participants = n_distinct(participant_id),
    daily_avg = round(surveys / 7, 1)
  ) %>%
  arrange(desc(surveys))

# Add rank
facility_activity$rank <- seq_len(nrow(facility_activity))

kable(facility_activity,
      caption = "Facility Performance Rankings",
      col.names = c("Facility", "Surveys", "Participants", "Daily Avg", "Rank"))
```

## Facility Comparison

```{r, fig.cap="Survey Distribution by Facility"}
facility_activity %>%
  head(10) %>%
  ggplot(aes(x = reorder(meta_facility_name, surveys), y = surveys, fill = surveys)) +
  geom_bar(stat = "identity") +
  scale_fill_gradient(low = "#95a5a6", high = "#27ae60") +
  coord_flip() +
  theme_minimal() +
  labs(title = "Top 10 Facilities by Survey Count",
       x = "Facility",
       y = "Number of Surveys") +
  guides(fill = "none") +
  theme(text = element_text(size = 12))
```

# Hourly Patterns

## Survey Timing Analysis

```{r, fig.cap="Survey Completions by Hour of Day"}
week_data %>%
  mutate(hour = hour(meta_completed_at)) %>%
  group_by(hour) %>%
  summarise(count = n()) %>%
  ggplot(aes(x = hour, y = count, fill = count)) +
  geom_bar(stat = "identity") +
  scale_fill_gradient(low = "#ecf0f1", high = "#3498db") +
  scale_x_continuous(breaks = seq(0, 23, 2)) +
  theme_minimal() +
  labs(title = "Survey Completions by Hour of Day",
       subtitle = "Peak hours for data collection",
       x = "Hour of Day (24-hour format)",
       y = "Number of Surveys") +
  guides(fill = "none") +
  theme(text = element_text(size = 12))
```

# Data Quality

## Response Completeness

```{r}
completeness_summary <- week_data_long %>%
  group_by(variable) %>%
  summarise(
    total = n(),
    complete = sum(!is.na(numeric_value) | !is.na(text_value)),
    completeness = round((complete / total) * 100, 1)
  ) %>%
  arrange(completeness)

# Categorize completeness
completeness_categories <- data.frame(
  Category = c("Excellent (95-100%)", "Good (80-94%)", "Fair (60-79%)", "Poor (<60%)"),
  Count = c(
    sum(completeness_summary$completeness >= 95),
    sum(completeness_summary$completeness >= 80 & completeness_summary$completeness < 95),
    sum(completeness_summary$completeness >= 60 & completeness_summary$completeness < 80),
    sum(completeness_summary$completeness < 60)
  )
)

kable(completeness_categories,
      caption = "Data Quality Distribution",
      col.names = c("Completeness Category", "Number of Variables"))
```

## Variables with Low Completion

```{r}
low_completion <- completeness_summary %>%
  filter(completeness < 80) %>%
  head(10)

if (nrow(low_completion) > 0) {
  kable(low_completion,
        caption = "Variables Requiring Attention (< 80% Complete)",
        col.names = c("Variable", "Total Responses", "Complete", "Completeness %"))
} else {
  cat("All variables have completion rates above 80%")
}
```

# Recruitment Patterns

## Coupon Usage

```{r}
if ("coupon_presented" %in% names(week_data_long)) {
  coupon_data <- week_data_long %>%
    filter(variable == "coupon_presented") %>%
    group_by(text_value) %>%
    summarise(count = n()) %>%
    arrange(desc(count))

  if (nrow(coupon_data) > 0) {
    kable(head(coupon_data, 10),
          caption = "Top 10 Recruiting Coupons",
          col.names = c("Coupon ID", "Recruits"))
  }
}
```

## Recruitment Chain Analysis

```{r, fig.cap="Daily New vs Returning Participants"}
if ("participant_type" %in% names(week_data_long)) {
  week_data %>%
    mutate(date = as.Date(meta_completed_at)) %>%
    group_by(date) %>%
    summarise(
      new_participants = sum(participant_type == "new", na.rm = TRUE),
      returning = sum(participant_type == "returning", na.rm = TRUE)
    ) %>%
    pivot_longer(cols = c(new_participants, returning), names_to = "type", values_to = "count") %>%
    ggplot(aes(x = date, y = count, fill = type)) +
    geom_bar(stat = "identity", position = "dodge") +
    scale_fill_manual(values = c("new_participants" = "#3498db", "returning" = "#e74c3c")) +
    theme_minimal() +
    labs(title = "New vs Returning Participants",
         x = "Date",
         y = "Number of Participants",
         fill = "Type") +
    theme(text = element_text(size = 12))
}
```

# Key Insights

## Performance Highlights

```{r}
# Find best performing day
best_day <- daily_completions %>%
  filter(surveys == max(surveys))

# Find most active facility
top_facility <- facility_activity %>%
  filter(rank == 1)

insights <- c(
  paste("• Best performing day:", best_day$weekday[1], format(best_day$date[1], "%B %d"), "with", best_day$surveys[1], "surveys"),
  paste("• Top facility:", top_facility$meta_facility_name[1], "completed", top_facility$surveys[1], "surveys"),
  paste("• Average daily completion rate:", round(total_surveys / 7, 1), "surveys"),
  paste("• Data quality:", round(mean(completeness_summary$completeness), 1), "% average completeness")
)

cat(paste(insights, collapse = "\n"))
```

## Recommendations

```{r}
# Generate recommendations based on data
recommendations <- character()

# Check for low activity days
low_days <- daily_completions %>% filter(surveys < mean(surveys) * 0.5)
if (nrow(low_days) > 0) {
  recommendations <- c(recommendations,
    paste("• Consider investigating low activity on:", paste(low_days$weekday, collapse = ", ")))
}

# Check for facilities with low performance
low_facilities <- facility_activity %>% filter(surveys < 5)
if (nrow(low_facilities) > 0) {
  recommendations <- c(recommendations,
    paste("• Support needed for", nrow(low_facilities), "facilities with fewer than 5 surveys"))
}

# Check for data quality issues
quality_issues <- sum(completeness_summary$completeness < 80)
if (quality_issues > 0) {
  recommendations <- c(recommendations,
    paste("•", quality_issues, "variables have completion rates below 80%"))
}

if (length(recommendations) > 0) {
  cat(paste(recommendations, collapse = "\n"))
} else {
  cat("• All metrics are within acceptable ranges\n• Continue current operational procedures")
}
```

---

*Report generated on `r format(Sys.Date(), "%B %d, %Y")` at `r format(Sys.time(), "%I:%M %p")`*
*Data period: `r format(week_start, "%B %d, %Y")` to `r format(week_end, "%B %d, %Y")`*